---
title: "DS 202 - lab #5: Drunk Driving and Fatal Accidents "
author: "Yumou Qiu"
ratio: 16x10
output:
  rmdshower::shower_presentation:
    self_contained: false
    katex: true
    theme: ribbon
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice #5: Accident data


## Outline

- Find your team members
- Create an R markdown file that contains all code to read the data and create the plots
- Homework 10 submit!


## FARS data

- US Department of transportation is keeping a record of every accident that results in a fatality in the FARS Data base (fatal accident report system, http://www.nhtsa.gov/FARS)
- FARS consists of 20+ tables consisting of various aspects of each accident
Documentation at
- https://www-fars.nhtsa.dot.gov/Main/index.aspx
- three of the main tables are [accident](https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/accident.csv), [person](https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/person.csv), and [vehicle](https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/vehicle.csv), stored in the [data foler](https://github.com/yumouqiu/DS202-Spring2023/tree/main/Practice/data/fars2016) of our Github page.

## Data Background (2)

- Data of the three main tables is made available from our website.
- Load the accident data, plot latitude and longitude on a map.
- Variable names of the `accident.csv`:

```{r, echo = FALSE}
acc <- read.csv("data/fars2016/accident.csv", stringsAsFactors = FALSE)
names(acc)
```

## Part one: Accident data

- are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?
- what time of the day do accidents happen (use variable HOUR)?
- what is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?


## Part two: Connecting data

- Connect to the [person table](https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/person.csv). Identify drivers (PER_TYP == 1, see fars manual ) and subset on them.
- Join accident and driver table (work out which variable(s) to use)
- Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!


## Submission

In Canvas, submit an RMarkdown file (and the knitted html document) with: 

  a. the names of all team members that worked on the document.
  b. a link to the github repo you were working in.
  c. code and answers to all of the questions.

