---
title: "DS 202 - Working in teams: Sample solution for the test case"
author: "Yumou Qiu"
ratio: 16x10
output:
  rmdshower::shower_presentation:
    self_contained: false
    katex: true
    theme: ribbon
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=4)
library(tidyverse)
```

## Sample Solution to the test case 

## A test case

We are asked to : 

1. Visualize life expectancy over time for Canada in the 1950s and 1960s using a line plot.
2. Something is clearly wrong with this plot! Turns out there's a data error 
in the data file: life expectancy for Canada in the year 1957 is coded
as `999999`, it should actually be `69.96`. Make this correction.
3. Visualize life expectancy over time for Canada again, with the corrected data.

Using the gapminder data provided

## Step-by-step plan of work

1. Read the data
2. Focus on values of Canada
3. Visualize the values for Canada
4. See whether we can find the problem
5. ... then we'll see ...

## Reading the data 

```{r}
life5060 <- read.csv("https://raw.githubusercontent.com/ds202-at-isu/materials/master/01_collaborative-environment/data/gapminder-5060.csv")
head(life5060)
dim(life5060)
```

## Focus on the values for Canada

```{r}
canada <- life5060 %>% filter(country == "Canada")
head(canada)
dim(canada)
```


##

- %>% means pass argument to a function

```{r}
log(exp(abs(-3)))
-3 %>% abs() %>% exp() %>% log()
```


## Another way of data processing

```{r}
canada1 <- life5060[life5060$country == "Canada", ]
head(canada1)
dim(canada1)
```

## Visualize 

Draw a line for the life expectancy in Canada
```{r}
canada %>% 
  ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

## Find the problematic value and fix it

```{r}
canada %>% filter(year == 1957)

canada_fixed <- canada %>% mutate(
  lifeExp = replace(lifeExp, year==1957, 69.96)  
)

canada_fixed
```

##

- Another way

```{r}
canada_fixed1 = canada1
canada_fixed1$lifeExp[canada_fixed1$year==1957] = 69.96
canada_fixed1
```

## visualize the fixed data

```{r}
canada_fixed %>% ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

## Use basic plot function in r

```{r}
plot(x = canada_fixed$year, y = canada_fixed$lifeExp, 
     type = 'l', xlab = "year", ylab = "lifeExp")
```

## ... on to the stretch goal ... 

Add lines for Mexico and the US

```{r}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, group=country)) + geom_line() +
  geom_line(data = life5060 %>% 
              filter(country %in% c("United States", "Mexico")))
```


## Use basic plot function in r

```{r}
plot(x = canada_fixed$year, y = canada_fixed$lifeExp, 
     type = 'l', xlab = "year", ylab = "lifeExp")

US = life5060[life5060$country == "United States", ]
Mexico = life5060[life5060$country == "Mexico", ]

lines(x = US$year, y = US$lifeExp)
lines(x = Mexico$year, y = Mexico$lifeExp)
```

## ... on to the stretch goal ... 

Add lines for Mexico and the US, color lines by country and add a legend

```{r, fig.width = 6}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, colour=country)) + geom_line() +
  geom_line(data = life5060 %>% 
              filter(country %in% c("United States", "Mexico")))

```

## How reproducible is this solution?

1. Navigate to https://yumouqiu.github.io/DS202-Spring2022/
2. Open the folder `01_collaborative-environment` and download the file `02_test-case-solution.Rmd`
3. Open the file in RStudio.
4. "Knit" the file.


## RMarkdown

### This is a RMarkdown presentation

- code and text/documentation are interwoven: reproducible and self-documenting.
- extend or refine analyses by copying and modifying code blocks.
- disseminate your work by sharing the RMarkdown file
