---
title: "DS 202: dplyr functions"
author: Yumou Qiu
ratio: 16x10
output:
  rmdshower::shower_presentation:
    self_contained: false
    katex: true
    theme: ribbon
---

## Messy (2)

## separate-1

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
df <- data.frame(x = c("a.b", "a.d", "b.c", NA))
df
df %>% separate(x, into = c("A", "B"))
```

## separate-2

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
df <- data.frame(x = c("a.b", "a,d", "b c", NA))
df
df %>% separate(x, into = c("A", "B"))
```

## separate-3

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
df <- data.frame(x = c("a.b", "a.d", "bc", NA))
df
df %>% separate(x, into = c("A", "B"))
```

## separate-4

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
df <- data.frame(x = c("a.b.c", "a.d; bc", "bc;a", NA))
df
df %>% separate(x, into = c("A", "B", "C"))
```

## Your Turn (5 min)


The Iowa Data Portal is a wealth of information on and about the State of Iowa. 

The website 
[Campaign Expenditures](https://data.iowa.gov/Government/Campaign-Expenditures/3adi-mht4/data) provides data on campaign expenditures. The code below reads the data into an R session.


```
url <- "https://data.iowa.gov/api/views/3adi-mht4/rows.csv"
campaign <- readr::read_csv(url)
```

Assess the 'messiness' of the data. List issues that prevent us from working with the data directly. Which of these issues are of type (1) or (2) of messiness?

## Problems with the data

- `Date` is text, in the format of Month/Day/Year (Messy 2)
- city coordinates are a combination of City name, state, zip code and geographic latitude and longitude. (Messy 2)
<!--- - expenditure amount is a textual expression, not a number (Messy different) --->
- Check the help for the function `parse_number` in the `readr` package and use it on the campaign expenditure data if there is a dollar sign.

no Messy 1? - problems of type Messy 1 are typically hard to detect and often up to interpretation/dependent on the analysis to be done.


## Your Turn (10 min)

- Use `separate` to split the date variable into year, month and day.
- Use `separate` again to extract geographic latitude and longitude (Hint - you might need several steps of `separate`)
- Using `dplyr` functionality answer the following questions:

  1. What is the total amount of campaign expenditure?
  2. What is the single largest cost in the campaign expenditures?
  3. What time frame does the campaign expenditure cover?

```{r, echo = FALSE, eval = FALSE}
campaign = read.csv("data/Iowa_Campaign_Expenditures.csv")
campaign1 = campaign %>% separate(Date, into = c("month", "day", "year"))

campaign11 = campaign1 %>% separate('City Coordinates', into = c("City1", "State1", "Zip1", "Latitude", "Longitude"))
campaign12 = campaign1 %>% separate('City Coordinates', into = c("Address", "Location"), sep = "(")
campaign13 = campaign1 %>% separate('City Coordinates', into = c("Address", "Location"), sep = "\n")
campaign14 = campaign1 %>% separate('City Coordinates', into = c("Address", "Location"), sep = "\\(")
campaign15 = campaign1 %>% separate('City Coordinates', into = c("Address", "Location"), sep = "\n\\(")

campaign2 = campaign15 %>% separate(Location, into = c("Latitude", "Longitude"), sep = ",") %>% select(-Address)
campaign2 = campaign2 %>% separate(Longitude, into = c("Longitude", "drop"), sep = "\\)") %>% select(-drop)
#as.numeric(campaign2$Latitude)
#as.numeric(campaign2$Longitude)
campaign3 = campaign2 %>% mutate(Latitude = as.numeric(campaign2$Latitude), Longitude = as.numeric(campaign2$Longitude))

options(dplyr.width = Inf)

campaign3 %>% summarise(TotalExp = sum(`Expenditure Amount`, na.rm = TRUE), MaxExp = max(`Expenditure Amount`, na.rm = TRUE))
campaign3 %>% filter(`Expenditure Amount` == max(`Expenditure Amount`, na.rm = TRUE))
campaign3 %>% group_by(year, month) %>% summarise(TotalExp = sum(`Expenditure Amount`, na.rm = TRUE), MaxExp = max(`Expenditure Amount`, na.rm = TRUE))

as.numeric(campaign3$year) %>% summary()

campaign3 %>% ggplot(aes(x = Longitude, y = Latitude)) + geom_point()
campaign3 %>% ggplot(aes(x = Longitude, y = Latitude)) + geom_point(aes(color = log(`Expenditure Amount`)))

campaignState = campaign3 %>% count(State)
sum(campaign3$State == "IA", na.rm = TRUE) / dim(campaign3)[1]
```
