---
title: "Polishing plots"
author: Yumou Qiu
ratio: 16x10
output:
  rmdshower::shower_presentation:
    self_contained: false
    katex: true
    theme: ribbon
---

# Polishing plots

## Outline

- color choices
- themes

## `ggplot2` provides defaults ...

- but every aspect of the plot can be changed
- colors are controlled through **scales**
- **themes** control presentation of non-data elements

# Colors

## Color Scales

default continuous colour scheme

```{r, message = FALSE, fig.height =4}
library(tidyverse)
p1 <- mpg %>% filter(year == 2008) %>%
  ggplot(aes(x = cty, y = hwy, colour = cyl)) +
  geom_point()
```


```{r, fig.height =4}
p1 + scale_colour_continuous()
```

## Color Scales

default discrete colour scheme

```{r, message = FALSE, fig.height =4}
p2 <- mpg %>% filter(year == 2008) %>%
  ggplot(aes(x = cty, y = hwy, colour = factor(cyl))) +
  geom_point()
```


```{r, fig.height =4}
p2 + scale_colour_discrete()
```

## Color Scales

- Colors are controlled through scales
`scale_colour_discrete` (`scale_colour_hue`) and `scale_colour_continuous` (`scale_colour_gradient`) are the default choices for factor variables and numeric variables
- we can change parameters of the default scale, or we can change
the scale function

## Colour gradients

`scale_colour_gradient (..., low = "#132B43", high = "#56B1F7", space = "Lab", na.value = "grey50", guide = “colourbar")`

- colors can be specified by hex code, name or through rgb()
- gradient goes from low to high - that should match the interpretation of the mapped variable


## Colour gradients - divergent scheme

`scale_colour_gradient2(..., low = muted("red"), mid = "white",  
high = muted("blue"), midpoint = 0, space = "Lab", na.value = "grey50",  guide = "colourbar")`

- midpoint is value of the ‘neutral’ color gradient2 is a divergent color scheme
- best matches a variable that goes from large negative to zero to large positive (or below mean, above mean)


## Discrete color schemes

`scale_color_hue (..., h = c(0, 360) + 15, c = 100, l = 65, h.start = 0, direction = 1, na.value = "grey50")`

- uses hue, chroma and luminance (=value)
- each level of a variable is assigned a different level of h


## Discrete color schemes - Brewer

`scale_colour_brewer(..., type = "seq", palette = 1, direction = 1)`

- brewer schemes are defined in RColorBrewer (Neuwirth, 2014) palettes can be specified by name or index
- see also http://colorbrewer2.org/ (Brewer et al 2002)

## All brewer schemes

```{r, fig.height=6}
library(RColorBrewer)
display.brewer.all()
```

## Color and Fill

- Area geoms (barcharts, histograms, polygons) use `fill` to map values to the fill color
- continuous color scales: `scale_fill_gradient`, `scale_fill_gradient2`, ...
- discrete color scales:  `scale_fill_hue`, `scale_fill_brewer`, `scale_fill_grey`, ...


# Themes

## Themes

- Themes allow to control every aspect of non-data related aspects of a plot 
- Several pre-defined themes: `theme_grey` (default), `theme_bw`, `theme_light`, `theme_dark`
- Use `theme_set` if you want it to apply a theme to every future plot, e.g. `theme_set(theme_bw())`
- `ggthemes` package defines additional themes: 
`library(help = "ggthemes")` lists all themes

## Example

```{r}
p <- mpg %>% ggplot(aes(x = displ, y =  cty, colour= factor(class))) + geom_point()
```

<p class="double">
```{r, fig.width = 4, fig.height = 2}
p + theme_grey()
```
```{r, fig.width = 4, fig.height = 2}
p + theme_bw()
```
</p>

## Example - more themes

```{r}
p <- mtcars %>% ggplot(aes(x = wt, y =  mpg, colour= factor(cyl))) + geom_point()
```

<p class="double">
```{r, fig.width = 4, fig.height = 2}
p + theme_light()
```
```{r, fig.width = 4, fig.height = 2}
p + theme_dark()
```

</p>

## More themes

```{r}
library(ggthemes)
```

<p class="double">
```{r, fig.width = 4, fig.height = 2}
p + theme_excel() 
```
```{r, fig.width = 4, fig.height = 2}
p + theme_fivethirtyeight()
```

</p>

## Elements

- You can also make your own theme, or modify an existing.
- Themes are made up of elements which can be one of:
`element_line`,  `element_text`, `element_rect`,
`element_blank`
- Gives you a lot of control over plot appearance.

## Elements of themes

- Axis:  
  `axis.line`, `axis.text.x`, `axis.text.y`, `axis.ticks`, `axis.title.x`, `axis.title.y`
- Legend:  
`legend.background`, `legend.key`, `legend.text`, `legend.title`
- Panel:  
`panel.background`, `panel.border`, `panel.grid.major`, `panel.grid.minor`
- Strip (facetting):  
`strip.background`, `strip.text.x`, `strip.text.y`

for a complete overview see `?theme`


## Changing elements manually

- to change an element add the theme function and specify inside:
- example:

```{r, fig.width = 4, fig.height = 4}
mpg %>% ggplot(aes(x = manufacturer)) + geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))
```    
